I always told myself that I would write these acknowledgements only at the end of my academic journey.
 However, it seems I may never want to leave this world, 
 so the time has come to thank all the dear people who have been with me throughout this long journey. 
 I want to emphasize that whatever I am, have done, and will do, is all thanks to you: 
 I have merely done what needed to be done, and you have been my guide. 
 I am endlessly grateful to you all.

First and foremost, 
 I must thank my family for always standing by me, 
 even when it seemed like I was lost. 
 In particular, I want to thank my mom for teaching me what it means to love, 
 my dad for showing me the meaning of dedication and sacrifice, 
 and my grandparents—grandpa Giuseppe for his kindness and grandma Rosa for her boundless dedication to work, for the meals prepared, and for her zest for life. 
 A special thanks go to my older brother Cristiano; 
 one of the most brilliant people I know.
 He always encouraged me to do my best and to never give up, 
 he has been a shining star in my darker moments—you are the best, keep it up!

I would also like to thank my lifelong friends, 
 Giovanni, Cri, and Sonia. 
 They have seen so many versions of me and still accept me--I do not know how they manage! 
 You have been by my side in tough times, providing lightness, and for that, I sincerely thank you. 
 I hope you continue to tolerate my craziness.
%
A special thanks go to Mone and Pedro who made me feel at home and believed in me during my university years--you are like brothers to me! 
%
A note of gratitude also goes to Jo, 
 my roommate for these last few months, 
 for helping me through difficult times--I apologize for all the therapy sessions you have had to endure!
%
I would like to express my sincere gratitude to my colleagues in Area 4.0.
 Our exhilarating launches and absurd discussions have made difficult times easier to navigate. 
 I assure you that I will continue to make coffee, even when I sometimes lose my composure.
%
Last but not least, I want to thank Marta--my muse and my reference point for so many years. 
 Much of who I am today is thanks to her. 
 I am still not as radiant as I would like to be,
 but some of the light I do have comes from the years spent with her. 
 Thank you for everything.
%
A special thanks to my second family, the Luffarelli--thank you for loving me as a son. 
 You will forever be in my heart.
%

Turning to the scientific side, 
 I want to express my boundless gratitude to my three main mentors: proff. Mirko Viroli, Roberto Casadei, and Danilo Pianini. 
 To you, I want to say, ``having access to people smarter than yourself is a blessing.'' 
% 
Mirko exemplifies the kind of professor people aspire to be--his dedication, love for his work, 
 and ongoing enthusiasm have always pushed me to give my best, because ``there are no alternatives to excellence''. 
 Roberto has been a reference for my work since my bachelor's thesis, 
 he is for me like a lighthouse guiding me through stormy seas--his advice, work ethic, and endless abilities are something I have always aspired to. 
 A special thank you to Danilo, 
 who has given me so much not just technically and scientifically 
 but also encouraged me to be a better version of myself (in his atypical methods).
 Additionally, I extend another heartfelt thank you to him. The cohesive group of researchers in Cesena exists primarily due to his unconventional efforts (e.g., the mandatory Friday Sthop).
%
Special mentions go to Alessandro Ricci for his dedication to teaching and his endless curiosity, 
 and to Giovanni Ciatto, who despite his outward cynicism, shows rare love for his work and displays a humanity hard to find in academia. 
 My only regret is that he beats me badly at Mario Kart.
%
I would also like to thank my students, especially Davide, Angelo, Francesco, and Giacomo. 
 You have taught me much more than I have imparted to you--I hope I've shared some of my joy for this work and given you some valuable guidance.
%
A special note also goes to Nico, you have helped me a lot in this last period of my PhD.
Conversations with you consistently enrich my perspective and inspire positive change in my work and thinking! 
I hope to live up to being a good example for you, just as my mentors have been for me.
Give more value to the sentence: Conversations with you always leave me with something good

Further, thanks go to my first office colleagues, dott. Angelo Croatti and prof. Sara Montagna, who were my go-to people in the initial dark months related to COVID of my PhD at our PS Lab.
%
Thank you to prof. Lukas Esterle for hosting me in Aarhus and guiding me attentively. 
Moreover, during my time abroad, I felt at home thanks to all my dormitory colleagues. 
You made those three beautiful months better.
%
I would also like to extend my gratitude to the researchers with whom I have engaged in thoughtful discussions, including Stefano Mariani, Mirco Musolesi, Guido Salvaneschi, Andrea Omicini, Giorgio Audrito, Giovanni Delnevo, and Morten From Elvebakken. To those not explicitly mentioned, your time and insights have been invaluable in my growth.

In closing, 
 many friends with whom I have shared memorable moments have not been explicitly acknowledged. 
 To you--from my hometown, to those who were with me in the early stages of my academic journey, and to those who have stood by me even in the final tough months of my doctoral studies--each of you has left an indelible imprint on my life. 
 I earnestly hope that I have been able to contribute something meaningful to your lives, just as you have enriched mine.